Mallows, Black-box combinatorial optimization, limited budget, ???

### 2020-10-22 paper/main.tex
more comments

\usepackage{amsmath,amssymb}
%\usepackage[square,numbers,sort&compress]{natbib}
\usepackage{xspace}
%% Useful packages
\usepackage{graphicx}
\usepackage{pdflscape,adjustbox}
\usepackage{multirow}
\usepackage[]{algorithm2e}
\usepackage{booktabs} % For formal tables
\usepackage{enumitem}
\usepackage{afterpage,tabularx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage[square,numbers,sort]{natbib}
% Fixes for using natbib and llncs
\makeatletter
% required for natbib to have "References" printed and as section*, not chapter*
\renewcommand\bibsection%
{\section*{\refname\@mkboth{\MakeUppercase{\refname}}{\MakeUppercase{\refname}}}}
\renewcommand\@biblabel[1]{#1.}
\def\bibfont{\small}
\makeatother

@@

\DeclareMathOperator*{\argmax}{arg\,max}

@@

@@

\subsection{Distributions over permutations}
Distributions over permutations~\cite{critchlow91} are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$. One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations. The MM defines the probability of each permutation $\sigma$ as follows:
%
\begin{equation}
\Prob(\sigma)=\frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi}
\end{equation}
%
with two parameters, $\theta$ and $\sigma_0$, \manuel{Permutation $\sigma_0$ a reference}{where $\sigma_0$ is the}  permutation that has the largest probability value, i.e., the mode of the distribution. The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay. The distance $d(\sigma,\sigma_0)$ is the Kendall's $\tau$ distance. The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.

@@

@@

Essentially, the algorithm  considers a sample of permutations along with the fitness of each of them. It is an iterative process in which the number of iterations is fixed. 
Then, at each iteration, UMM (i) estimates a surrogate distribution of the fitness function and (ii) proposes a good fitted individual.

\section{Unbalanced Mallows model: probabilistic BB optimization}
In this section we describe our main contribution: a novel algorithm for the optimization of a black box function over permutation spaces. This algorithm is called Unbalanced Mallows model (UMM) since it a probabilistic method based on a well-known probability distribution for permutations: the Mallows model.

\begin{algorithm}[t]
 \KwData{$n$, $m$,}
 \KwResult{how to write algorithm with \LaTeX2e }
 $S$\;
 $F$\;
 $\hat\sigma_0$ = uBorda($S, F,\rho$)\;
 \For{$m$ \text{iterations}}{
  linear decrease on $\theta$\;
  $\sigma$ = sample from a $MM(\hat\sigma_0, \theta)$\;
  Append $\sigma$ and $f(\sigma)$ to $S$ and $F$\;
   $\hat\sigma_0$ = uBorda($S, F,\rho$)\;
 }
 \Return $S, F$
 \caption{Unbalanced Mallows model for black box optimization}
 \label{alg:umm}
\end{algorithm}

The high-level behavior of UMM is shown in Algorithm~\ref{alg:umm}. It starts with an \textit{small} sample of $m'$ permutations, $S=\{\sigma_1, \ldots,  \sigma_{m'}\}$ that are randomly generated. UMM also keeps the function evaluation of each solution, $F=\{f(\sigma_1), \ldots, f(\sigma_{m'})\}$. An probabilistic model is used internally to guide the search. Then, for $m$ iterations\EKHINE{esto es exactamente así o es $m-m'$?}, UMM generates a random, promising candidate solution $\sigma$ and includes it in the sample $S$ while $f(\sigma)$ in included in $F$. At each iteration, the internal probabilistic model is updated.

>>>>

\usepackage{amsmath,amssymb}
%\usepackage[square,numbers,sort&compress]{natbib}
\usepackage{xspace}
%% Useful packages
\usepackage{graphicx}
\usepackage{pdflscape,adjustbox}
\usepackage{multirow}

\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{booktabs} % For formal tables
\usepackage{enumitem}
\usepackage{afterpage,tabularx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage[square,numbers,sort]{natbib}
% Fixes for using natbib and llncs
\makeatletter
% required for natbib to have "References" printed and as section*, not chapter*
\renewcommand\bibsection%
{\section*{\refname\@mkboth{\MakeUppercase{\refname}}{\MakeUppercase{\refname}}}}
\renewcommand\@biblabel[1]{#1.}
\def\bibfont{\small}
\makeatother

@@

\newcommand{\true}{\textsc{true}}
\newcommand{\false}{\textsc{false}}
\newcommand{\NULL}{\textsc{null}}
\newcommand{\NA}{---}
\newcommand{\assign}{\ensuremath{:=}}
\newcommand{\procedure}[1]{\textsf{#1}}
\newcommand{\Inf}{\ensuremath{\infty}}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

Let us assume a black-box ``fitness'' function that must be minimized over a
space of permutations $f\colon S_n \to \mathbb{R}$. Being black-box means that
we can evaluate any candidate permutation $\sigma \in S_n$ to obtain
$f(\sigma)$, however, we do not know anything else about the form of
$f$. Moreover, the evaluation of $f$ is expensive in computation time or
resources and, thus, limits the budget of candidate permutations that we can
evaluate to a relatively low number. Here, we will study budgets lower than
400. Such expensive black-box combinatorial problems arise in
manufacturing~\citep{FerAlvDiaIglEna2014ants}, where the evaluation function
may involve expensive simulations for which no closed-form mathematical
description is available.

@@

\subsection{Distributions over permutations}
Distributions over permutations~\cite{critchlow91} are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$. One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations. The MM defines the probability of each permutation $\sigma$ as follows:
%
\begin{equation}
\Prob(\sigma)=\frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi}
\end{equation}
%
with two parameters, $\theta$ and $\sigma_0$, \manuel{Permutation $\sigma_0$ a reference}{where $\sigma_0$ is the}  permutation that has the largest probability value, i.e., the mode of the distribution. The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay. The distance $d(\sigma,\sigma_0)$ is the Kendall's $\tau$ distance.\MANUEL{Can we define briefly Kendall's tau?} The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.

@@

\section{Unbalanced Mallows model: probabilistic BB optimization}

@@

Essentially, the algorithm  considers a sample of permutations along with the fitness of each of them. It is an iterative process in which the number of iterations is fixed. 
Then, at each iteration, UMM (i) estimates a surrogate distribution of the fitness function and (ii) proposes a good fitted individual. \MANUEL{Hay un poco de duplicación entre lo anterior y lo siguiente, no?}

In this section we describe our main contribution: a novel algorithm for the optimization of a black box function over permutation spaces. This algorithm is called Unbalanced Mallows model (UMM) since it a probabilistic method based on a well-known probability distribution for permutations: the Mallows model.\MANUEL{this sounds as if this was the first time you talk about MM, but you already described it above, no? }

\begin{algorithm}[t]
 \caption{UMM: Unbalanced Mallows Model for Black-Box Optimization}
 \label{alg:umm}
  \begin{algorithmic}[1]
    \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
    \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
    \State linear decrease on $\theta$
    \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
    \State $S \assign S \cup \{\sigma\}$
    \State $F \assign F \cup \{f(\sigma)\}$
    \EndFor 
    \State \Return $S$, $F$
  \end{algorithmic}
\end{algorithm}

As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $m'$ permutations,\MANUEL{but the algorithm requires $m$, no $m'$?} $S=\{\sigma_1, \dotsc,  \sigma_{m'}\}$ that are randomly generated. UMM also keeps the function evaluation of each solution, $F=\{f(\sigma_1), \ldots, f(\sigma_{m'})\}$.  Then, for $m$ iterations\EKHINE{esto es exactamente así o es $m-m'$?}, UMM generates a random, promising candidate solution $\sigma$ from a probabilistic model and includes it in the sample $S$ while $f(\sigma)$ is included in $F$. At each iteration, the internal probabilistic model is updated.
